---
title: "occCite: Downloading and Citing Biodiversity Data"
author:
- Hannah L. Owens
- Cory Merow
- Brian Maitner
- Jamie M. Kass
- Vijay Barve
- Robert Guralnick
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
library(ape)
library(occCite)
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_knit$set(root.dir = paste0(path.package("occCite"), "/extdata/"))
```

<!-- badges: start -->
[![R build status](https://github.com/hannahlowens/occCite/workflows/R-CMD-check/badge.svg)](https://github.com/hannahlowens/occCite/actions)
[![cran version](https://www.r-pkg.org/badges/version/occCite)](https://cran.r-project.org/package=occCite)
[![rstudio mirror downloads](https://cranlogs.r-pkg.org/badges/occCite)](https://github.com/r-hub/cranlogs.app)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Codecov test coverage](https://codecov.io/gh/hannahlowens/occCite/branch/main/graph/badge.svg)](https://codecov.io/gh/hannahlowens/occCite?branch=main)
[![](https://badges.ropensci.org/407_status.svg)](https://github.com/ropensci/software-review/issues/407)
<!-- badges: end -->

Please cite rgbif. Run the following to get the appropriate citation for the version you're using:

```r
citation(package = "rgbif")
```

To get started, see:

* occCite introduction vignette (https://docs.ropensci.org/rgbif/articles/rgbif.html)
* occCite advanced feature vignette 
* Function reference (https://docs.ropensci.org/rgbif/reference/index.html): an overview of all `rgbif` functions.
* Articles (https://docs.ropensci.org/rgbif/articles/index.html): vignettes/tutorials on how to download data, clean data, and work with taxonomic names.

Check out the `rgbif` [paper][] for more information on this package and the sister [Python][pygbif], [Ruby][gbifrb], and [PHP][phpgbif] clients.

Note: Maximum number of records you can get with `occ_search()` and `occ_data()` is 100,000. See https://www.gbif.org/developer/occurrence

## Installation

```{r eval=FALSE}
install.packages("rgbif")
```

Or, install development version

```{r eval=FALSE}
pak::pkg_install("ropensci/rgbif")
# OR
install.packages("rgbif", repos="https://dev.ropensci.org")
```

```{r}
library("rgbif")
```


## Meta

* Please [report any issues or bugs](https://github.com/hannahlowens/occCite/issues).
* License: GPL-3
* Get citation information for `occCite` in R using `citation(package = 'occCite')`

- - -

For an overview tutorial video of the package, see [our YouTube video](https://www.youtube.com/watch?v=7qSCULN_VjY&t=17s).
