---
title: "Advanced search and citation of occurrences"
author:
- Hannah L. Owens
- Cory Merow
- Brian Maitner
- Jamie M. Kass
- Vijay Barve
- Robert Guralnick
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Advanced execution of occurrence searches and citations generation}
  \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(ape)
library(occCite)
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_knit$set(root.dir = system.file('extdata/', package='occCite'))
```

# Advanced features

This vignette demonstrates more advanced features and customization available in `occCite`. We recommend you read `vignette("Simple.Rmd", package = "occCite")` first, if you have not already done so.

## Loading data from previous GBIF searches

Querying GBIF can take quite a bit of time, especially for multiple species and/or well-known species. In this case, you may wish to access previously-downloaded data sets from your computer by specifying the general location of your downloaded `.zip` files. `occQuery` will crawl through your specified `GBIFDownloadDirectory` to collect all the `.zip` files contained in that folder and its subfolders. It will then import the most recent downloads that match your taxon list. These GBIF data will be appended to a BIEN search the same as if you do the simple real-time search (if you chose BIEN as well as GBIF), as was shown above. `checkPreviousGBIFDownload` is `TRUE` by default, but if `loadLocalGBIFDownload` is `TRUE`, `occQuery` will ignore `checkPreviousDownload`. It is also worth noting that `occCite` does not currently support mixed data download sources. That is, you cannot do GBIF queries for some taxa, download previously-prepared data sets for others, and load the rest from local data sets on your computer.

```{r simple_search, eval=F}
# Simple search
myOldOccCiteObject <- occQuery(x = "Protea cynaroides",
                                  datasources = c("gbif", "bien"),
                                  GBIFLogin = GBIFLogin, 
                                  GBIFDownloadDirectory = 
                                    system.file('extdata/', package='occCite'),
                                  checkPreviousGBIFDownload = T)
```

```{r simple_search sssssecret cooking show, eval=T, echo = F}
# Simple search
data(myOccCiteObject)
myOldOccCiteObject <- myOccCiteObject
```

Here is the result. Look familiar?

```{r simple_search_loaded_GBIF_results}
#GBIF search results
head(myOldOccCiteObject@occResults$`Protea cynaroides`$GBIF$OccurrenceTable);
#The full summary
summary(myOldOccCiteObject)
```

Getting citation data works the exact same way with previously-downloaded data as it does from a fresh data set.

```{r getting_citations_from_already-downloaded_GBIF_data}
#Get citations
myOldOccCitations <- occCitation(myOldOccCiteObject)
print(myOldOccCitations)
```

Note that you can also load multiple species using either a vector of species names or a phylogeny (provided you have previously downloaded data for all of the species of interest), and you can load occurrences from non-GBIF data sources (e.g. BIEN) in the same query.

***

## Performing a Multi-Species Search

In addition to doing a simple, single species search, you can also use `occCite` to search for and manage occurrence datasets for multiple species. You can either submit a vector of species names, or you can submit a *phylogeny*! The occCitation function will return a named list of citation tables in the case of multiple species.

## occCite with a Phylogeny

Here is an example of how such a search is structured, using an unpublished phylogeny of billfishes.

```{r multispecies_search_with_phylogeny, eval=T, echo=T}
library(ape)
#Get tree
treeFile <- system.file("extdata/Fish_12Tax_time_calibrated.tre", package='occCite')
phylogeny <- ape::read.nexus(treeFile)
tree <- ape::extract.clade(phylogeny, 22)
#Query databases for names
myPhyOccCiteObject <- studyTaxonList(x = tree, 
                                     datasources = "GBIF Backbone Taxonomy")
#Query GBIF for occurrence data
myPhyOccCiteObject <- occQuery(x = myPhyOccCiteObject, 
                            datasources = "gbif",
                            GBIFDownloadDirectory = system.file('extdata/', package='occCite'),
                            loadLocalGBIFDownload = T,
                            checkPreviousGBIFDownload = F)
# What does a multispecies query look like?
summary(myPhyOccCiteObject)
```

When you have results for multiple species, as in this case, you can also plot the summary figures either for the whole search...

```{r plotting all species, eval=T, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide', fig.hold='hold', out.width="100%"}
plot(myPhyOccCiteObject)
```

*or* you can plot the results by species!

```{r plotting phylogenetic search by species, eval=T, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide', fig.hold='hold', out.width="100%"}
plot(myPhyOccCiteObject, bySpecies = T, plotTypes = c("yearHistogram", "source"))
```

And then you can print out the citations, separated by species (or not, but in this example, they're separate).

```{r getting_citations_for_a_multispecies_search, echo=T}
#Get citations
myPhyOccCitations <- occCitation(myPhyOccCiteObject)

#Print citations as text with accession dates.
print(myPhyOccCitations, bySpecies = T)
```
