---
title: "occCite Vignette: Downloading and Citing Biodiversity Data"
author:
- Hannah Owens
- Cory Merow
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{An overview of executing searches and generating citations using occCite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(occCite)
library(ape)
```

# Introduction 

We have entered the age of data-intensive scientific discovery. As datasets increase in complexity and heterogeneity, we must preserve the cycle of data citation from primary data sources to aggregating databases to research products and back to primary data sources. The citation cycle keeps science transparent, but it is also key to supporting primary providers by documenting the use of their data. The Global Biodiversity Information Facility and other data aggregators have made great strides in harvesting citation data from research products and linking them back to primary data providers. However, this only works if those that publish research products cite primary data sources in the first place. We developed `occCite`, a set of `R`-based tools for downloading, managing, and citing biodiversity data, to advance toward the goal of closing the data provenance cycle. These tools preserve links between occurrence data and primary providers once researchers download aggregated data, and facilitate the citation of primary data providers in research papers.

The `occCite` work flow follows a three-step process. First, the user inputs one or more taxonomic names (or a phylogeny). `occCite` then rectifies those names by checking them against one or more taxonomic databases, which can be specified by the user (see the \href{http://gni.globalnames.org/}{Global Names List}). The results of the taxonomic rectification are then kept in an `occCiteData` object in local memory. Next, `occCite` takes the `occCiteData` object and user-defined search parameters to query BIEN (through `rbien`) and/or GBIF(through `rGBIF`) for records. The results are appended to the `occCiteData` object, along with metadata on the search. Finally, the user can pass the `occCiteData` object to `occCitation`, which compiles citations for the primary providers, database aggregators, and `R` packages used to build the dataset.

Future iterations of `occCite` will track citation data through the data cleaning process and provide a series of visualizations on raw query results and final datasets. It will also provide data citations in a format congruent with best-practice recommendations for large biodiversity datasets.

# Setup

If you plan to query GBIF, you will need to provide them with your user login information. We have provided a dummy login below to show you the format. *You will need to provide actual account information.* This is because you will actually be downloading *all* of the records available for the species using `occ_download()`, instead of getting results from `occ_search()`, which has a hard limit of 200,000 occurrences.

```{r login, eval=FALSE}
library(occCite);

#Creating a GBIF login
GBIFLogin <- GBIFLoginManager(user = "occCiteTester",
                          email = "****@yahoo.com",
                          pwd = "12345");
```

# Performing a simple search

## The basics

At its simplest, `occCite` allows you to search for occurrences for a single species. The taxonomy of the user-specified species will be verified using EOL and NCBI taxonomies by default. 

```{r simple_search, eval=FALSE}
# Simple search
mySimpleOccCiteObject <- occQuery(x = "Protea cynaroides",
                            datasources = c("gbif", "bien"),
                            GBIFLogin = GBIFLogin, 
                            GBIFDownloadDirectory = system.file("extdata", package = "occCite"),
                            checkPreviousGBIFDownload = T);
```

It is also possible to load data from a previously-downloaded GBIF dataset by specifying the location of the `.zip` file. These GBIF data will be appended to a BIEN search the same as if you do the simple real-time search, as shown above. Note that `checkPreviousGBIFDownload` is `TRUE` by default, but if `loadLocalGBIFDownload` is `TRUE`, `occQuery` will ignore `checkPreviousDownload`.

```{r simple_load}
# Simple load
mySimpleOccCiteObject <- occQuery(x = "Protea cynaroides", datasources = c("gbif", "bien"), 
                                  GBIFLogin = NULL, 
                                  GBIFDownloadDirectory = system.file("extdata", package = "occCite"),
                                  loadLocalGBIFDownload = T,
                                  checkPreviousGBIFDownload = F);
```

```{r simple_search_GBIF_results}
# GBIF search results
head(mySimpleOccCiteObject@occResults$`Protea cynaroides`$GBIF$OccurrenceTable);
```

```{r simple_search_BIEN_results}
#BIEN search results
head(mySimpleOccCiteObject@occResults$`Protea cynaroides`$BIEN$OccurrenceTable);
```

There is also a summary method for `occCite` objects with some basic information about your search.

```{r summary of simple search}
summary(mySimpleOccCiteObject)
```

## Simple citations

After doing a search for occurrence points, you can use `occCitation()` to generate citations for primary biodiversity databases, as well as database aggregators.
**Note:** Currently, GBIF and BIEN are the only aggregators for which citations are supported.

```{r simple_citation}
#Get citations
mySimpleOccCitations <- occCitation(mySimpleOccCiteObject);
```
```{r show_simple_citations}
cat(paste(mySimpleOccCitations$Citation, sep = ""), sep = "\n");
```

## Simple Taxonomic Rectification

In the simplest of searches, such as the one above, the taxonomy of your input species name is automatically rectified through the `occCite` function `studyTaxonList()` using `gnr_resolve()` from the `taxize` `R` package. If you would like to change the source of the taxonomy being used to rectify your species names, you can specify as many taxonomic repositories as you like from the Global Names Index (GNI). The complete list of GNI repositories can be found [here](http://gni.globalnames.org/data_sources). 

`studyTaxonList()` chooses the taxonomic names closest to those being input and documents which taxonomic repositories agreed with those names. `studyTaxonList()` instantiates an `occCiteData` object the same way `occQuery()` does. This object can be passed into `occQuery()` to perform your occurrence data search.

```{r taxonomic_rectification}
#Rectify taxonomy
myTROccCiteObject <- studyTaxonList(x = "Protea cynaroides", 
                                  datasources = c("NCBI", "EOL", "ITIS"));
myTROccCiteObject@cleanedTaxonomy
```

***
# Loading data from previous GBIF searches

Querying GBIF can take quite a bit of time, especially for multiple species and/or well-known species. In this case, you may wish to access previously-downloaded datasets from your computer. Here is a simple example of how that is done.

```{r search_using_previously-downloaded_GBIF_data}
#Query databases for names
myOldOccCiteObject <- studyTaxonList(x = "Protea cynaroides", datasources = "NCBI");

#Access GBIF data from a specified download directory
##Note: you do not need a login for this.
myOldOccCiteObject <- occQuery(x = myOldOccCiteObject, 
                            datasources = "gbif",
                            GBIFDownloadDirectory = system.file("extdata/", package = "occCite"),
                            loadLocalGBIFDownload = T);
```

Here is the result. Look familiar?

```{r simple_search_loaded_GBIF_results}
#GBIF search results
head(myOldOccCiteObject@occResults$`Protea cynaroides`$GBIF$OccurrenceTable);

#The full summary
summary(myOldOccCiteObject)
```

Getting citation data works the exact same way with previously-downloaded data as it does from a fresh dataset.

```{r getting_citations_from_already-downloaded_GBIF_data}
#Get citations
myOldOccCitations <- occCitation(myOldOccCiteObject);
cat(paste(myOldOccCitations$Citation, sep = ""), sep = "\n");
```

Note that you can also load multiple species using either a vector of species names or a phylogeny (provided you have previously downloaded data for all of the species of interest), and you can load occurrences from non-GBIF datasources (e.g. BIEN) in the same query.

***

# Performing a Multi-Species Search

In addition to doing a simple, single species search, you can also use `occCite` to search for and manage occurrence datasets for multiple species. You can either submit a vector of species names, or you can submit a *phylogeny*!

## occCite with a Phylogeny

Here is an example of how such a search is structured, using an unpublished phylogeny of billfishes.

```{r multispecies_search_with_phylogeny, eval=T, echo=T}
library(ape);

#Get tree
treeFile <- system.file("extdata/Fish_12Tax_time_calibrated.tre", package='occCite');
tree <- ape::read.nexus(treeFile);

#Query databases for names
myPhyOccCiteObject <- studyTaxonList(x = tree, datasources = "NCBI");

#Query GBIF for occurrence data
myPhyOccCiteObject <- occQuery(x = myPhyOccCiteObject, 
                            datasources = "gbif",
                            GBIFDownloadDirectory = system.file("extdata/", package = 'occCite'), 
                            loadLocalGBIFDownload = T,
                            checkPreviousGBIFDownload = F);

```

```{r getting_citations_for_a_multispecies_search, echo=T}
# What does a multispecies query look like?
summary(myPhyOccCiteObject)

#Get citations
myPhyOccCitations <- occCitation(myPhyOccCiteObject);
cat(paste(myPhyOccCitations$Citation, sep = ""), sep = "\n");
```
