
---
title: "occCite Vignette: Downloading and Citing Biodiversity Data"
author:
- Hannah Owens
- Cory Merow
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{An overview of making rangeModelMetadata objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

# Introduction 

*occCite* is an `R`-based application for downloading, managing, and citing biodiversity data. This package is designed to facilitate querying multiple biodiversity database aggregators, tracking sources through the occurrence point cleaning process, and citing primary datasources, as well as database aggregator services.

# Setup

If you plan to query GBIF, you will need to provide them with your user login information. We have provided a dummy login below. You will need to provide actual account information.

```{r login}
library(occCite);

#Creating a GBIF login
login <- GBIFLoginManager(user = "wallacetester",
                          email = "cmerow@yahoo.com",
                          pwd = "wallacetester");
```

# Performing a Simple Search

## The Basics

At its simplest, *occCite* allows you to search for occurrences for a single species. The taxonomy of the user-specified species will be verified using EOL and NCBI taxonomies.

```{r simple search}
##Simple search
myOccCiteObject <- occQuery(x = "Protea cynaroides",
                            datasources = c("gbif", "bien"),
                            GBIFLogin = login, 
                            GBIFDownloadDirectory = "~/Desktop");
```

```{r simple search GBIF results}
#GBIF search results
head(myOccCiteObject@occResults$`Protea cynaroides`$GBIF$OccurrenceTable);
```

```{r simple search BIEN results}
#GBIF search results
head(myOccCiteObject@occResults$`Protea cynaroides`$BIEN$OccurrenceTable);
```

## Simple Taxonomic Rectification

The taxonomy of your input species name is rectified using `gnr_resolve()` from the `taxize` `R` package. If you would like to change the source of the taxonomy being used to rectify your species names, you can specify as many taxonomic repositories as you like from the Global Names Index (GNI). The complete list of GNI repositories can be found [here](http://gni.globalnames.org/data_sources). `taxonRectification()` is an internal function in `studyTaxonList()` in `occCite` that chooses the taxonomic names closest to those being input and documents which taxonomic repositories agreed with those names.

```{r taxonomic rectification}
#Rectify taxonomy
myOccCiteObject <- studyTaxonList(x = "Protea cynaroides", 
                                  datasources = c("NCBI", "EOL", "ITIS"));
myOccCiteObject@cleanedTaxonomy
```

## Simple Citations

After doing a search for occurrence points, you can use `occCitation()` to generate citations for primary biodiversity databases, as well as database aggregators.
**Note:** Currently, GBIF is the only aggregator for which citations are supported.

```{r simple citation}
#Do a simple search
myOccCiteObject <- studyTaxonList(x = "Protea cynaroides", 
                                  datasources = c("NCBI", "EOL", "ITIS"));
myOccCiteObject <- occQuery(x = myOccCiteObject, 
                            GBIFLogin = login, 
                            GBIFDownloadDirectory = "~/Desktop");

#Get citations
myOccCitations <- occCitation(myOccCiteObject);
```
```{r show simple citations}
cat(paste(myOccCitations$Citation, sep = ""), sep = "\n");
```

***

# Performing a Multi-Species Search

In addition to doing a simple, single species search, you can also use *occCite* to search for and manage occurrence datasets for multiple species. You can either submit a vector of species names, or you can submit a *phylogeny!*

## occCite with a Phylogeny

Here is an example of a search done using an unpublished phylogeny of billfishes.

```{r multispecies search with phylogeny}
library(ape);

#Get tree
# try
treeFile <- system.file("extdata/Fish_12Tax_time_calibrated.tre", package='occCite');
tree <- read.nexus(treeFile);

#Query databases for names
myOccCiteObject <- studyTaxonList(x = tree, datasources = "NCBI");

#Query GBIF for occurrence data
login <- GBIFLoginManager(user = "wallacetester",
                          email = "cmerow@yahoo.com",
                          pwd = "wallacetester");
myOccCiteObject <- occQuery(x = myOccCiteObject, 
                            GBIFLogin = login, 
                            datasources = "gbif",
                            GBIFDownloadDirectory = "~/Desktop");
```

```{r getting citations for a multispecies search}
#Get citations
myOccCitations <- occCitation(myOccCiteObject);
cat(paste(myOccCitations$Citation, sep = ""), sep = "\n");
```
